\documentclass[11pt]{article}

\usepackage[T1]{fontenc, lucidbry}
\usepackage{mynotes,cite}

\pagestyle{notes}
\renewcommand{\NotesTitle}{FK Dimensionless Variables, }

\begin{document}

\section{Dimensionless FK Model}

The Hamiltonian for the basic Frenkel-Kontorova model is
\begin{equation}
H=\sum_{i=1}^N\left[\frac{{p_i}^2}{2m_i}
	+\frac{g_i}{2}\left(z_i-z_{i-1}-a_i\right)^2
	+\frac{V_0}{2}\left(1-\cos\frac{2\pi z_i}{\lambda}\right)\right].
\end{equation}
Some of the notation here and in the following is taken from the review paper by Braun and Kivshar on the FK equation, \textit{Physics Reports} \textbf{306}, pp. 1--108 (1998). The parameters that appear are $m_i$, the mass of particle $i$, $g_i$ and $a_i$, the strength and equilibrium length of the spring that connects particle $i$ to particle $i-1$, $V_0$, the peak-to-peak amplitude of the substrate potential, and $\lambda$, the wavelength of the substrate potential.

The equations of motion are
\begin{align}
\dot{z}_i&=\frac{\d H}{\d p_i}=\frac{p_i}{m_i},\\
\dot{p}_i&=-\frac{\d H}{\d z_i}=-g_i(z_i-z_{i-1}-a_i)
	+g_{i+1}(z_{i+1}-z_i-a_{i+1})
	-\frac{\pi V_0}{\lambda}\sin\frac{2\pi z_i}{\lambda}.
\end{align}
We'll work with dimensionless variables. To start, we'll define the average mass through
\begin{equation}
m=\frac{1}{N}\sum_{i=1}^N m_i.
\end{equation}
To nondimensionalize, define the following constants:
\begin{equation}
t_0 = \frac{\lambda}{2\pi}\sqrt{\frac{2m}{V_0}},\quad
p_0 = \sqrt{\frac{mV_0}{2}},\quad
g_0 = \frac{2\pi^2V_0}{\lambda^2}.
\end{equation}
Then define the following dimensionless variables
\begin{equation}
\tau=\frac{t}{t_0},\quad
\phi_i=\frac{2\pi}{\lambda} z_i,\quad
\rho_i=\frac{p_i}{p_0},\quad
\alpha_i=\frac{2\pi}{\lambda} a_i,\quad
\mu_i=\frac{m_i}{m},\quad
\gamma_i=\frac{g_i}{g_0}.
\end{equation}
In terms of these variables, the equations of motion become
\begin{align}
\phi_i'&=\frac{\rho_i}{\mu_i},\\
\rho_i'&=-\gamma_i(\phi_i-\phi_{i-1}-\alpha_i)
	+\gamma_{i+1}(\phi_{i+1}-\phi_i-\alpha_{i+1})
	-\sin\phi_i.
\end{align}
Note that if all the masses and springs are identical, then $\mu_i=1$ and all the $\alpha_i$ are equal so that they cancel, and this reduces to a one-parameter set of equations.

We can add a damping force $-\eta p_i$ and a constant external driving force, $f_i$, to the original equations, to obtain
\begin{align}
\dot{z}_i&=\frac{p_i}{m_i},\\
\dot{p}_i&=-g_i(z_i-z_{i-1}-a_i)
	+g_{i+1}(z_{i+1}-z_i-a_{i+1})
	-\eta p_i+f_i
	-\frac{\pi V_0}{\lambda}\sin\frac{2\pi z_i}{\lambda}.
\end{align}
Defining the variables
\begin{equation}
\beta=t_0\eta,\quad
\epsilon_i = \frac{t_0}{p_0}f_i,
\end{equation}
the dimensionless version of the equations of motion become
\begin{align}
\phi_i'&=\frac{\rho_i}{\mu_i},\\
\rho_i'&=-\gamma_i(\phi_i-\phi_{i-1}-\alpha_i)
	+\gamma_{i+1}(\phi_{i+1}-\phi_i-\alpha_{i+1})
	-\beta\rho_i+\epsilon_i
	-\sin\phi_i.
\end{align}
If we now go back and express the original Hamiltonian in terms of the dimensionless variables, we have
\begin{equation}
H=\frac{V_0}{2}\sum_{i=1}^N
	\left[\frac{{\rho_i}^2}{2\mu_i}
	+\frac{\gamma_i}{2}\left(\phi_i-\phi_{i-1}-\alpha_i\right)^2
	+\left(1-\cos\phi_i\right)\right].
\end{equation}
In summary, the three important parameters are the dimensionless spring constant,
\begin{equation}
\gamma_i=\frac{g_i}{g_0}=\frac{\lambda^2}{2\pi^2V_0}\,g_i,
\end{equation}
the dimensionless damping,
\begin{equation}
\beta=t_0\eta=\frac{\lambda}{2\pi}\sqrt{\frac{2m}{V_0}}\,\eta,
\end{equation}
and the dimensionless external forcing,
\begin{equation}
\epsilon_i=\frac{t_0}{p_0}\,f_i=\frac{\lambda}{\pi V_0}\,f_i.
\end{equation}

\end{document}
